---
title: "PS270 Final Project"
author: "Fayzaan Virk"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

```{r, include=FALSE}
library(tidyverse)
library(usdata)
```


## Project thoughts

I'm not really sure what to do for the final project or what the scope of it is. I did have an idea about the mayoral race in Minneapolis. One of the candidates, Omar Fateh, was Somali, but I saw somewhere that only about half of the Somali population in Minneapolis voted for him and the rest voted for the incumbent Jacob Frey. I wanted to see if I could find data for this and why he split his own community's vote.

## My Research Question

Unfortunately, I was not able to find any data for my above thoughts about the Minneapolis election. I decided to transition to another question relating to a statement I have often heard: "Asian-Americans are the model minority". I wonder if there is any statistical backing to this. We can already see from the census data I imported that Asian-Americans tend to have higher than average educational-attainment and career earnings, but does this translate to their communities. I intend to use the county_complete data set to examine if a higher proportion of Asians in a community will lead to a higher median income. My hypothesis is that they will. Because of their individual education and earnings, I believe the presence of Asian-Americans in a community will economically benefit that community. My explanatory variable will be the percentage of the population in a county that is Asian. If there's time, I can also examine other economic or social metrics like unemployment rate, percentage under the poverty line, crime rate, and life expectancy. My outcome variable will be the county's median household income. A positive regression coefficient would support my hypothesis, while one that is negative or close to zero would show no association or a negative association between the presence of Asians and median income.

```{r}
data(county_complete)
asian <- read_csv("data/asian_breakdown.csv")
education <- read_csv("data/educational_attainment.csv")
```

I checked the output for head() and glimpse() and they both work. The reason I did not include them is because the tibble is 188 columns and would take up way too much space when I knit.
