---
title: "PS270 Final Project"
author: "Fayzaan Virk"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

```{r, include=FALSE}
library(tidyverse)
library(usdata)
```


## Project thoughts

I'm not really sure what to do for the final project or what the scope of it is. I did have an idea about the mayoral race in Minneapolis. One of the candidates, Omar Fateh, was Somali, but I saw somewhere that only about half of the Somali population in Minneapolis voted for him and the rest voted for the incumbent Jacob Frey. I wanted to see if I could find data for this and why he split his own community's vote.

## My Research Question

Unfortunately, I was not able to find any data for my above thoughts about the Minneapolis election. I decided to transition to another question relating to a statement I have often heard: "Asian-Americans are the model minority". I wonder if there is any statistical backing to this. We can already see from the census data I imported that Asian-Americans tend to have higher than average educational-attainment and career earnings, but does this translate to their communities. I intend to use the county_complete data set to examine if a higher proportion of Asians in a community will lead to a higher median income. My hypothesis is that they will. Because of their individual education and earnings, I believe the presence of Asian-Americans in a community will economically benefit that community. My explanatory variable will be the percentage of the population in a county that is Asian. My outcome variable will be the county's median household income. If there's time, I can also examine other economic or social metrics like unemployment rate, percentage under the poverty line, crime rate, and life expectancy. A positive regression coefficient would support my hypothesis, while one that is negative or close to zero would show no association or a negative association between the presence of Asians and median income.

```{r}
data(county_complete)
asian <- read_csv("data/asian_breakdown.csv")
education <- read_csv("data/educational_attainment.csv")
```

I checked the output for head() and glimpse() and they both work. The reason I did not include them is because the tibble is 188 columns and would take up way too much space when I knit.

## Data Wrangling and Visualization

For my visualization, I want to quickly show the relationship between proportion of Asian American population and median income. While it may look like this is almost the end of answering the question I had, I still have many confounding variables to tackle to more strongly establish a causatory relationship. I'm going to create a scatterplot with Asian proportion on the x-axis and median individual income on the y-axis. The county_complete data set already has these variables neatly stored, so I don't have to do any wrangling to get the plot. However, since the assignment specifically requests I do some cleaning, I'm going to plot an average line and calculate the coefficient of determination to help see if there is a causatory relationship between my two variables.

First, I'm going to calculate the median individual income for all counties.

```{r}
county_complete |>
  summarize(med_inc = median(median_individual_income_2019))
```

Now, I will make a scatterplot with a smoothed line and the median line.

```{r}
ggplot(county_complete, aes(x = asian_2017, y = median_individual_income_2019)) + 
  geom_point() + 
  scale_x_log10() + 
  geom_smooth() + 
  labs(
    x = "Proportion of Asian Americans",
    y = "Median Individual Income in Dollars",
    title = "Median Income and Asian Population by County"
  ) + 
  geom_hline(yintercept = 26975, color = "red",  lwd = 1)
```

Now, I will find a linear model for the line of best fit to see the coefficient of determination.

```{r}
prop_med_fit <- lm(median_individual_income_2019 ~ asian_2017, county_complete)
prop_med_fit
prop_med_fit_summary = summary(prop_med_fit)
prop_med_fit_summary$r.squared
```




